Place of death in NRS COVID-19 deaths

### 2021-04-14 covid_nrs_place_of_death.Rmd
Commented out bit of code where I was trying to cluster case rates & home deaths - not used at the minute

```{r}
cumulative_case_2020_based_grouping_la <-
  weekly_covid_cases_la %>% 
  filter(ref_area!="Scotland") %>%
  filter(year == 2020) %>% 
  filter(week_number == max(week_number)) %>%  # keep last week
  select(ref_area, cumulative_positive, crude_rate_positive_per100k)

cumulative_case_2020_based_grouping_hb <-
  weekly_covid_cases_hb %>% filter(year == 2020) %>% 
  filter(ref_area!="Scotland") %>%
  filter(year == 2020) %>% 
  filter(week_number == max(week_number)) %>%  # keep last week 
  select(ref_area, cumulative_positive, crude_rate_positive_per100k)

k_means_clustering_cases_and_per100k_2020_hb <- map(
  .x = 2:4,
  .f = ~kmeans(x=cumulative_case_2020_based_grouping_hb %>% select(cumulative_positive, crude_rate_positive_per100k), centers = .x, nstart = 1e3)
) %>% set_names(nm=paste0("k",2:4))

clusters_hb <-
  bind_cols(
    cumulative_case_2020_based_grouping_hb %>% select(ref_area),
    map_dfc(.x = k_means_clustering_cases_and_per100k_2020_hb, .f = function(x) return(x$cluster))
  )

cumulative_case_2020_based_grouping_hb <-
  cumulative_case_2020_based_grouping_hb %>%
  left_join(clusters_hb, by="ref_area") %>%
  mutate(
    quartiles_cases = ntile(cumulative_positive, n=4),
    quartiles_per100k = ntile(crude_rate_positive_per100k, n=4)
  )

# TODO: apply to LA also!

hb_cases_plots <- list()

hb_cases_plots$cumulative_cases_vs_per100k <-
  cumulative_case_2020_based_grouping_hb %>%
  ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area)) +
  geom_point() +
  geom_label_repel() +
  # theme_minimal() +
  labs(
    x="COVID cases",
    y="COVID cases per 100,000 population",
    title="Cases vs. population rate"
  ) +
  NULL

hb_cases_plots$cumulative_cases_vs_per100k_log10 <-
  cumulative_case_2020_based_grouping_hb %>%
  ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area)) +
  geom_point() +
  geom_label_repel() +
  # theme_minimal() +
  labs(
    x="COVID cases (log)",
    y="COVID cases per 100,000 population (log)",
    title="Cases vs. population rate (log scale axes)"
  ) +
  scale_x_log10() +
  scale_y_log10() +
  NULL

hb_cases_plots$cumulative_cases_vs_per100k_quartiles_cases <-
  cumulative_case_2020_based_grouping_hb %>%
  ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(quartiles_cases))) +
  geom_point(size=3, aes(colour=factor(quartiles_cases))) +
  scale_colour_viridis_d() +
  geom_label_repel() +
  # theme_minimal() +
  labs(
    x="COVID cases",
    y="COVID cases per 100,000 population",
    title="Cases vs. population rate"
  ) +
  NULL

hb_cases_plots$cumulative_cases_vs_per100k_quartiles_cases <-
  cumulative_case_2020_based_grouping_hb %>%
  ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(quartiles_per100k))) +
  geom_point(size=3, aes(colour=factor(quartiles_cases))) +
  scale_colour_viridis_d() +
  geom_label_repel() +
  # theme_minimal() +
  labs(
    x="COVID cases",
    y="COVID cases per 100,000 population",
    title="Cases vs. population rate",
    colour = "Quartile (Cases per 100,000)"
  ) +
  NULL

hb_cases_plots$cumulative_cases_vs_per100k_k2 <-
  cumulative_case_2020_based_grouping_hb %>%
  ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(k2))) +
  geom_point(size=3, aes(colour=factor(k2))) +
  scale_colour_viridis_d() +
  geom_label_repel() +
  # theme_minimal() +
  labs(
    x="COVID cases",
    y="COVID cases per 100,000 population",
    title="Cases vs. population rate"
  ) +
  NULL

hb_cases_plots$cumulative_cases_vs_per100k_k3 <-
  cumulative_case_2020_based_grouping_hb %>%
  ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(k3))) +
  geom_point(size=3, aes(colour=factor(k3))) +
  scale_colour_viridis_d() +
  geom_label_repel() +
  # theme_minimal() +
  labs(
    x="COVID cases",
    y="COVID cases per 100,000 population",
    title="Cases vs. population rate"
  ) +
  NULL

hb_cases_plots$cumulative_cases_vs_per100k_k4 <-
  cumulative_case_2020_based_grouping_hb %>%
  ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(k4))) +
  geom_point(size=3, aes(colour=factor(k4))) +
  scale_colour_viridis_d() +
  geom_label_repel() +
  # theme_minimal() +
  labs(
    x="COVID cases",
    y="COVID cases per 100,000 population",
    title="Cases vs. population rate"
  ) +
  NULL

hb_cases_plots_assembled <- patchwork::wrap_plots(hb_cases_plots, ncol = 2)

hb_cases_plots_assembled
```

@@

```{r}
(figure_annual_deaths_by_cause <-
  total_deaths_cause_of_death %>%
  filter(year == 2020 & cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
  # pivot_longer(cols = c(deaths, annual_deaths_mean), names_to = "deaths_type", values_to = "deaths") %>%
  ggplot() +
  # geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths, group = deaths_type, linetype = deaths_type), position = "dodge", colour = "grey10") +
  geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths), position = "dodge", colour = "grey10") +
  geom_point(aes(x = cause_of_death, y = annual_deaths_mean), position = "dodge", colour = "grey10") +
  scale_fill_viridis_d() +
  facet_wrap(~place_of_death) +
  NULL
)
```

>>>>

NOTE: I've commented this out - I ended up not using it. TODO: remove in a future iteration!

```{r}
# cumulative_case_2020_based_grouping_la <-
#   weekly_covid_cases_la %>% 
#   filter(ref_area!="Scotland") %>%
#   filter(year == 2020) %>% 
#   filter(week_number == max(week_number)) %>%  # keep last week
#   select(ref_area, cumulative_positive, crude_rate_positive_per100k)
# 
# cumulative_case_2020_based_grouping_hb <-
#   weekly_covid_cases_hb %>% filter(year == 2020) %>% 
#   filter(ref_area!="Scotland") %>%
#   filter(year == 2020) %>% 
#   filter(week_number == max(week_number)) %>%  # keep last week 
#   select(ref_area, cumulative_positive, crude_rate_positive_per100k)
# 
# k_means_clustering_cases_and_per100k_2020_hb <- map(
#   .x = 2:4,
#   .f = ~kmeans(x=cumulative_case_2020_based_grouping_hb %>% select(cumulative_positive, crude_rate_positive_per100k), centers = .x, nstart = 1e3)
# ) %>% set_names(nm=paste0("k",2:4))
# 
# clusters_hb <-
#   bind_cols(
#     cumulative_case_2020_based_grouping_hb %>% select(ref_area),
#     map_dfc(.x = k_means_clustering_cases_and_per100k_2020_hb, .f = function(x) return(x$cluster))
#   )
# 
# cumulative_case_2020_based_grouping_hb <-
#   cumulative_case_2020_based_grouping_hb %>%
#   left_join(clusters_hb, by="ref_area") %>%
#   mutate(
#     quartiles_cases = ntile(cumulative_positive, n=4),
#     quartiles_per100k = ntile(crude_rate_positive_per100k, n=4)
#   )
# 
# # TODO: apply to LA also!
# 
# 
# hb_cases_plots <- list()
# 
# hb_cases_plots$cumulative_cases_vs_per100k <-
#   cumulative_case_2020_based_grouping_hb %>%
#   ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area)) +
#   geom_point() +
#   geom_label_repel() +
#   # theme_minimal() +
#   labs(
#     x="COVID cases",
#     y="COVID cases per 100,000 population",
#     title="Cases vs. population rate"
#   ) +
#   NULL
# 
# hb_cases_plots$cumulative_cases_vs_per100k_log10 <-
#   cumulative_case_2020_based_grouping_hb %>%
#   ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area)) +
#   geom_point() +
#   geom_label_repel() +
#   # theme_minimal() +
#   labs(
#     x="COVID cases (log)",
#     y="COVID cases per 100,000 population (log)",
#     title="Cases vs. population rate (log scale axes)"
#   ) +
#   scale_x_log10() +
#   scale_y_log10() +
#   NULL
# 
# hb_cases_plots$cumulative_cases_vs_per100k_quartiles_cases <-
#   cumulative_case_2020_based_grouping_hb %>%
#   ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(quartiles_cases))) +
#   geom_point(size=3, aes(colour=factor(quartiles_cases))) +
#   scale_colour_viridis_d() +
#   geom_label_repel() +
#   # theme_minimal() +
#   labs(
#     x="COVID cases",
#     y="COVID cases per 100,000 population",
#     title="Cases vs. population rate"
#   ) +
#   NULL
# 
# hb_cases_plots$cumulative_cases_vs_per100k_quartiles_cases <-
#   cumulative_case_2020_based_grouping_hb %>%
#   ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(quartiles_per100k))) +
#   geom_point(size=3, aes(colour=factor(quartiles_cases))) +
#   scale_colour_viridis_d() +
#   geom_label_repel() +
#   # theme_minimal() +
#   labs(
#     x="COVID cases",
#     y="COVID cases per 100,000 population",
#     title="Cases vs. population rate",
#     colour = "Quartile (Cases per 100,000)"
#   ) +
#   NULL
# 
# hb_cases_plots$cumulative_cases_vs_per100k_k2 <-
#   cumulative_case_2020_based_grouping_hb %>%
#   ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(k2))) +
#   geom_point(size=3, aes(colour=factor(k2))) +
#   scale_colour_viridis_d() +
#   geom_label_repel() +
#   # theme_minimal() +
#   labs(
#     x="COVID cases",
#     y="COVID cases per 100,000 population",
#     title="Cases vs. population rate"
#   ) +
#   NULL
# 
# 
# hb_cases_plots$cumulative_cases_vs_per100k_k3 <-
#   cumulative_case_2020_based_grouping_hb %>%
#   ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(k3))) +
#   geom_point(size=3, aes(colour=factor(k3))) +
#   scale_colour_viridis_d() +
#   geom_label_repel() +
#   # theme_minimal() +
#   labs(
#     x="COVID cases",
#     y="COVID cases per 100,000 population",
#     title="Cases vs. population rate"
#   ) +
#   NULL
# 
# 
# hb_cases_plots$cumulative_cases_vs_per100k_k4 <-
#   cumulative_case_2020_based_grouping_hb %>%
#   ggplot(data=., aes(x=cumulative_positive, y=crude_rate_positive_per100k, label=ref_area, shape=factor(k4))) +
#   geom_point(size=3, aes(colour=factor(k4))) +
#   scale_colour_viridis_d() +
#   geom_label_repel() +
#   # theme_minimal() +
#   labs(
#     x="COVID cases",
#     y="COVID cases per 100,000 population",
#     title="Cases vs. population rate"
#   ) +
#   NULL
# 
# 
# hb_cases_plots_assembled <- patchwork::wrap_plots(hb_cases_plots, ncol = 2)
# 
# hb_cases_plots_assembled

@@

```{r}
(figure_annual_deaths_by_cause <-
  total_deaths_cause_of_death %>%
  filter(year == 2020 & cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
  # pivot_longer(cols = c(deaths, annual_deaths_mean), names_to = "deaths_type", values_to = "deaths") %>%
  ggplot() +
  # geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths, group = deaths_type, linetype = deaths_type), position = "dodge", colour = "grey10") +
  geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths), colour = "grey10") +
  # geom_point(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  geom_line(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  scale_fill_viridis_d() +
  facet_wrap(~place_of_death) +
  NULL
)
```
